---
title: "Mini Assignment Mod 8 - Insurace Cost Model - Linear Regression"
output:
  html_document:
    df_print: paged
---

Kaggle Notebook: https://www.kaggle.com/shikhar4128/insurance-cost-model-using-linear-regression/notebook 
Data source: https://github.com/stedy/Machine-Learning-with-R-datasets 


```{r}
library(tidyverse) # metapackage with lots of helpful functions
library(psych) #used for describeBy function
library(scales) #used to make y axis in comma format
library(GGally) #used for ggcorr
library(ggcorrplot) #for ggcorrplot
```

```{r}
insu <- read.csv("insurance.csv", stringsAsFactors = T)
head(insu,10)
str(insu)
```

```{r}
summary(insu)
```

From summary(insu) we can derive:

1) Gender is evenly distributed
2) Min amd Max age in dataset is 18 and 64
3) Smoker to non smoker ration is almost 4:1
4) All 4 regions have equal datasets
5) Average Medical cost is $13,270.

# Creating yes/no categories
```{r}
insu$group <- ifelse(insu$charges > mean(insu$charges) ,"high","low") #create a col indicating charges are high or low
insu$isObese <- ifelse(insu$bmi >= 30,"yes","no") #Create a col indicating if person is classified as obese or not
```

# Evaluating other variables 
```{r}
describeBy(insu$charges,insu$region) #charges by region
```

```{r}
#charges by region
insu %>%
  group_by(region) %>%
  summarize(meanCharges = mean(charges)) %>%
  ungroup()
```

```{r}
#total amount charged by region
region_chrg <- insu %>% group_by (region) %>% summarize(Total_Charge =sum(charges)) %>%
                  arrange(desc(Total_Charge))
region_chrg
```

## EDA - Vizualization 
```{r}
#total charges by region
ggplot(data=insu,aes(x=reorder(region,-charges),y=charges)) + # - is used for sorting in decending order 
geom_bar(stat="identity",fill="steel blue") +
scale_y_continuous(labels = comma) +
labs(title="Total Charges By Region", x="Region",y="Charges")
```

```{r}
ggplot(data = insu,aes(region,charges)) + geom_boxplot(fill = c(2:5)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges per Region")
```

Based on the above chart we can say that region doesnt have a large impact on medical cost

```{r}
#charges by smoker classification
describeBy(insu$charges,insu$smoker)

```

```{r}
smoker_chrg <- insu %>% group_by(smoker) %>% summarize(Total_Charge=sum(charges)) %>%
                  arrange(desc(Total_Charge))
smoker_chrg
```

```{r}
ggplot(data=insu,aes(x=reorder(smoker,-charges),y=charges)) +
geom_bar(stat="identity",fill="steel blue") +
scale_y_continuous(labels = comma) +
labs(title="Smoker Vs Charges", x="Smoker",y="Charges")
```
 
```{r}
ggplot(data = insu,aes(smoker,charges)) + geom_boxplot(fill = c(2:3)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges by Smoking Status")
```
 
 Bases on the above boxplot we can say smokers spend more on medical cost.
 
 
```{r}
#charges by gender
describeBy(insu$charges,insu$sex)
```
 
```{r}
ggplot(data = insu,aes(sex,charges)) + geom_boxplot(fill = c(2:3)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges per Gender")
```

There is not that big of a difference between gender and medical charges, although males tend to have a wider interquartile range of charages. 

```{r}
#charges by children
describeBy(insu$charges,insu$children)
```

```{r}
ggplot(data = insu,aes(as.factor(children),charges)) + geom_boxplot(fill = c(2:7)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges per Children") +
xlab("Num of Children")
```

People with 2 children have more medical expenses. 5 children, have less expenses.

```{r}
#obesity vs charge
describeBy(insu$charges,insu$isObese)
```

```{r}
ggplot(data = insu,aes(as.factor(isObese),charges)) + geom_boxplot(fill = c(2:3)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges per Obesity") +
xlab("Obesity")
```
Customers classified as obese tend to incurr higher medical charges, although the median charge between the two does not differ largely.

```{r}
#Check the correlation between variables. We need to convert factor variables to numeric first before doing that.
insu$sex <- as.numeric(insu$sex)
insu$smoker <- as.numeric(insu$smoker)

#insu$sex <- NULL
#insu$region <- NULL #removing region and sex as it doesnt affect the charges as we have seen above in boxplot.

head(insu)
str(insu)

corr <- cor(insu[1:6]) #excluding group and obesity which we created.
corr
```

```{r}
ggcorrplot(corr) #by defaut square
ggcorrplot(corr,method="circle") #can do it in circle

corr

cor(insu[1:6])[,"charges"] #the value that most influence charges
ggcorrplot(corr,hc.order = TRUE, type = "lower",
   lab = TRUE)
```
 The three main factors effecting insurance charges (in decreasing order of importance) smoker, age, bmi (obesity)
 smoker =0.79 > age(0.29) > bmi (0.18)
 
# Regress charges by bmi with a scatter plot
```{r}
ggplot(insu, aes(x = bmi, y = charges)) +
  geom_point() +
  stat_smooth(method = 'lm')
```

```{r}
ggplot(insu, aes(x = age, y = charges)) +
  geom_point() +
  stat_smooth(method = 'lm')
```

 
# Model 1 - Regression Analysis: 
```{r}
lm1 <- lm(formula =charges ~ age + bmi +children +smoker ,data=insu)
summary(lm1)
```


# Model 2 - Linear Regression: Gives higher penalty to person who smokes and is obese

```{r}
lm2 <- lm (formula=charges ~ isObese * smoker ,data =insu)
summary(lm2)
```

#Residuals 
```{r}
resids <- insu %>%
  select(bmi, charges) %>%
  mutate(fittedCharges = -5409.2 + 13386.2*bmi
         , residuals = charges - fittedCharges)
head(resids)
```

```{r}
# Create a dataframe with residuals and identifying information
resids2 <- insu %>%
  select(age, bmi, children, smoker, charges)
resids2$fittedCharges = lm1$fitted.values
resids2$residuals = lm1$residuals

# Get the five best performing store/quarter combinations
best <- resids2 %>%
  arrange(desc(residuals)) %>%
  .[1:5,]
# Get the five worst performing store/quarter combinations
worst <- resids2 %>%
  arrange(residuals) %>%
  .[1:5,] %>%
  arrange(desc(residuals))

# Combine the five best and worst into one dataframe and display them
bestWorst <- bind_rows(best,worst)
bestWorst
```

# Model 3 - Linear Regression
Gives higher penalty to person who smokes and is older
```{r}
lm3 <- lm(formula=charges ~ age +smoker ,data=insu)
summary(lm3)
```

## Predicting Future Values
```{r}
# Dreate a dataframe of new observations
newObservations <- data.frame(age = c(25, 32, 43, 54, 65)
                              , bmi = c(19, 32, 43, 24, 15)
                              ,children = c(1, 2, 3, 4, 5)
                              , smoker = c(2, 1, 2, 1, 1))
# Add a new column of predicted values
newObservations$predictedCharges = predict(lm1, newObservations)
# Display the dataframe in this notebook
newObservations
```

```{r}
# Dreate a dataframe of new observations
newObservations_2 <- data.frame(age = c('25', '32', '43', '54', '65')
                              , isObese = c('no', 'yes', 'yes', 'no', 'no')
                              ,children = c(1, 2, 3, 4, 5)
                              , smoker = c(2, 1, 2, 1, 1))
# Add a new column of predicted values
newObservations_2$predictedCharges = predict(lm2, newObservations_2)
# Display the dataframe in this notebook
newObservations_2
```


***************************SUMMARY**********************
The business problem in this mini assignment to gauge the cost of insurance for individuals (employees) given their health qualifications. The analysis is predictive and nature and aimed at identifying the target variable, insurance price. The statistically significant predictor variables include age, smoking status, number of children, and obesity. Understanding these prediction cost can help an insurance company evaluate how much to charge those looking to get insurance or employers who will add employees onto a company health insurance plan. The linear regression algorithm was chosen as the data is used to make predictions and explain the relationships between the dependent variable (numerical NOT binary), and independent variables and to what degree of confidence we can attach to these predictions.

Model 1 Analysis: 
1) Smoking increases medical cost by USD 15059 
2) As the number of children increase the health cost increases by USD 431 
3) The older the person, the higher the medical cost

On this model we get a r-squared of 0.7489 which implies that 74% of variation of charges can be explained by the set of independent variables we have included. This model has all the independent variables available in the original dataset. Also all variables we have include are statistically significant predictors of charges (p value < 0.05 <- level of significance)

Model 2 Analysis: 
1) Obesity increase medical cost by $18,463 
2) Smoking increases medical cost by $13,386 
3) If you smoke and are classifyed as obese then medical cost is expected to increase by $19,329

Model 3 Analysis : 
1) As age increases the medical cost increases. 
2) Using just age and smoker as independent variables, we have built a model with an r-squared of 72.14% which is comparable with our first model which use all of the original variables.

Conclusion: 
1) I would use model 1 and also model 2 for penalty purpose. 
2) We can conclude smoking is the single most effective variable effecting charges